---
title: "Basic Regression"
author: "Corina Geier"
date: "2/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`r sixmonth_data = na.omit(read.csv(".\\prepped_data\\sixmonth.csv"))`

## Question 1

## Cases per capita with overall GHSI score

```{r}

summary(lm(casepc ~ overall, data = sixmonth_data))$coef

par(mfrow=c(2,2),mar=c(5,4,2,1))

plot(lm(casepc ~ overall, data = sixmonth_data))

```

Coefficient estimate: 0.0637

pvalue = 0.0387

## Deaths per capita with overall GHSI score


``` {r}

summary(lm(deathpc ~ overall, data = sixmonth_data))$coef

par(mfrow=c(2,2),mar=c(5,4,2,1))

plot(lm(deathpc ~ overall, data = sixmonth_data))

```

Coefficient estimate: 0.0046

pvalue < .0001

## Case Fatality Ratio with overall GHSI score 


```{r}

summary(lm(cfratio ~ overall, data = sixmonth_data))$coef

par(mfrow=c(2,2),mar=c(5,4,2,1))

plot(lm(cfratio ~ overall, data = sixmonth_data))

```

Coefficient estimate: 0.0556

pvalue = .0012


## Question 2

## Cases per capita with each subcomponent

```{r}

subcomponents <- c("prev_emergence_pathogens", "early_detection", "rapid_response", "robust_health_sector", "commitments", "risk_environment")

for(i in 1:length(subcomponents)) {
  predictor_i <- (subcomponents)[i]
  print(predictor_i)
  print(summary(lm(casepc ~ ., data = sixmonth_data[,c("casepc",predictor_i)]))$coef)
  par(mfrow=c(2,2),mar=c(5,4,2,1))
  plot(lm(casepc ~ ., data = sixmonth_data[,c("casepc",predictor_i)]))
}


```



## Deaths per capita with each subcomponent

```{r}
subcomponents <- c("prev_emergence_pathogens", "early_detection", "rapid_response", "robust_health_sector", "commitments", "risk_environment")

for(i in 1:length(subcomponents)) {
  predictor_i <- (subcomponents)[i]
  print(predictor_i)
  print(summary(lm(deathpc ~ ., data = sixmonth_data[,c("deathpc",predictor_i)]))$coef)
  par(mfrow=c(2,2),mar=c(5,4,2,1))
  plot(lm(deathpc ~ ., data = sixmonth_data[,c("deathpc",predictor_i)]))
}


```

## Case Fatality Ratio with each subcomponent

```{r}
subcomponents <- c("prev_emergence_pathogens", "early_detection", "rapid_response", "robust_health_sector", "commitments", "risk_environment")

for(i in 1:length(subcomponents)) {
  predictor_i <- (subcomponents)[i]
  print(predictor_i)
  print(summary(lm(cfratio ~ ., data = sixmonth_data[,c("cfratio",predictor_i)]))$coef)
  par(mfrow=c(2,2),mar=c(5,4,2,1))
  plot(lm(cfratio ~ ., data = sixmonth_data[,c("cfratio",predictor_i)]))
}


```

