---
title: "GDP_Analysis"
author: "Edward Lou"
date: "2/21/2021"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Imports & Pre-processing
```{r}
# import libraries
library(tidyverse)
library(lubridate)
library(ggpubr)
library(dplyr)

# load sixmonth data, omitting NA values
sixmonth_data = na.omit(read.csv(".\\prepped_data\\sixmonth.csv", check.names = FALSE))

# add gdp_pc quartile
sixmonth_data$gdp_pc_quartile <- ntile(sixmonth_data$gdp_pc, 4)
```
\newpage

# Correlation between GDP per-capita and overall GHSI index
```{r}
ggscatter(sixmonth_data,x='gdp_pc' ,y='overall',
          add='reg.line',cor.coef=TRUE,cor.method='pearson',
          xlab='GDP per-capita',ylab='Overall GHSI Score')
```
\newpage

# Baseline: Outcomes vs overall GHSI

## Cases per-capita vs overall GHSI
```{r}
summary(lm(casepc ~ overall, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(casepc ~ overall, data = sixmonth_data))
```
\newpage

## Deaths per-capita vs overall GHSI
```{r}
summary(lm(deathpc ~ overall, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(deathpc ~ overall, data = sixmonth_data))
```
\newpage

## Case-fatality ratio vs overall GHSI
```{r}
summary(lm(cfratio ~ overall, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(cfratio ~ overall, data = sixmonth_data))
```

\newpage

# Confounding GDP per-capita (quartiles): Outcomes vs overall GHSI

## Cases per-capita vs overall GHSI & GDP per-capita quartiles
```{r}
summary(lm(casepc ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(casepc ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))
```
\newpage

## Deaths per-capita vs overall GHSI & GDP per-capita quartiles
```{r}
summary(lm(deathpc ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(deathpc ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))
```
\newpage

## Case-fatality ratio vs overall GHSI & GDP per-capita quartiles
```{r}
summary(lm(cfratio ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(cfratio ~ overall + factor(gdp_pc_quartile), data = sixmonth_data))
```
\newpage

# Confounding GDP per-capita (raw): Outcomes vs overall GHSI

## Cases per-capita vs overall GHSI & GDP per-capita
```{r}
summary(lm(casepc ~ overall + gdp_pc, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(casepc ~ overall + gdp_pc, data = sixmonth_data))
```
\newpage

## Deaths per-capita vs overall GHSI & GDP per-capita
```{r}
summary(lm(deathpc ~ overall + gdp_pc, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(deathpc ~ overall + gdp_pc, data = sixmonth_data))
```
\newpage

## Case-fatality ratio vs overall GHSI & GDP per-capita
```{r}
summary(lm(cfratio ~ overall + gdp_pc, data = sixmonth_data))$coefficients
par(mfrow=c(2,2),mar=c(5,4,2,1))
plot(lm(cfratio ~ overall + gdp_pc, data = sixmonth_data))

```