---
title: "democracy_index_confounding"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of democracy index as a confounding variable

Start by loading the six month data:

```{r}
data <- read.csv(file = '../prepped_data/sixmonth.csv')
democracy <- read.csv(file = '../intermediate_data/democracy_index.csv')
democracy <- democracy[c("X2019", "country_code")]
colnames(democracy)[1] <- "democracy_index"
merged <- merge(data, democracy, by=c('country_code'), all.x=TRUE, all.y=FALSE)
```

### Regressions on democracy index

```{r}
summary(lm(formula = casepc ~ democracy_index, data = merged))
plot(merged$democracy_index, merged$casepc, pch=19)
```

```{r}
summary(lm(formula = deathpc ~ democracy_index, data = merged))
plot(merged$democracy_index, merged$deathpc, pch=19)
```

```{r}
summary(lm(formula = cfratio ~ democracy_index, data = merged))
plot(merged$democracy_index, merged$cfratio, pch=19)
```

From these plots, it appears that democracy index isn't a good predictor of
poor Covid-19 outcomes, but when you look at deaths-per-capita, there is a trend
where the highest deaths-per-capita are in high democracy index countries.

### Regressions on GHSI subcomponent scores and democracy index

```{r}
summary(lm(formula = casepc ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment, data = merged))
summary(lm(formula = casepc ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment + democracy_index, data = merged))
```

For this regression, we see that the democracy index coefficient is
statistically significant and that it has a negative relationship with
deaths-per-capita.