---
title: "democracy_index_confounding"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of democracy index as a confounding variable

Start by loading the six month data:

```{r}
data <- read.csv(file = '../prepped_data/six_month_outlier_screened.csv')
```

### Regressions on democracy index

```{r}
summary(lm(formula = cfratio ~ democracy_index, data = data))
plot(data$democracy_index, data$cfratio, pch=19)
```

From these plots, it appears that democracy index isn't a good predictor of
poor Covid-19 outcomes, but when you look at deaths-per-capita, there is a trend
where the highest deaths-per-capita are in high democracy index countries.

### Regressions on GHSI subcomponent scores and democracy index

```{r}
summary(lm(formula = casepc ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment, data = data))
summary(lm(formula = casepc ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment + democracy_index, data = data))
```

For this regression, we see that the democracy index coefficient is
statistically significant and that it has a negative relationship with
deaths-per-capita.