---
title: "island_confounding"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of island status as a confounding variable

Start by loading the six month data:

```{r}
data <- read.csv(file = '../prepped_data/six_month_outlier_screened.csv')
```


### Regressions on island nation status

Run regressions with is_island as the sole explanatory variable:

```{r}
summary(lm(formula = cfratio ~ factor(is_island), data = data))
plot(data$is_island, data$cfratio, pch=19)
```

Overall findings are that the island nation status alone is not a great
predictor of cases-per-capita, deaths-per-capita, and case fatality ratio, but
from the plots you can see that island nations are on average lower on all three
of these measures.

### Regressions on GHSI overall score and island nation status

```{r}
summary(lm(formula = cfratio ~ overall, data = data))
summary(lm(formula = cfratio ~ overall + factor(is_island), data = data))
plot(data$overall, data$cfratio, pch=19, col=as.factor(data$is_island))
```

In all three of these cases, adding is_island to the regression does not
meaningfully increase the R-squared measure or decrease the residual standard
error. So adding island nation status doesn't help explain changes in cases or
deaths relative to the GHSI scores.

### Regressions on GHSI subcomponent scores and island nation status

```{r}
summary(lm(formula = cfratio ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment, data = data))
summary(lm(formula = cfratio ~ prev_emergence_pathogens + early_detection + rapid_response + robust_health_sector + commitments + risk_environment + factor(is_island), data = data))
```

In this case adding the is_island variable improves the R-squared by about 10%,
but the overall R-squared and errors are pretty bad. It also should be noted
that is_island has a p-value of 0.08 which isn't statistically significant, but
is the 4th most significant of the 7 variables in the regression.